---
title: "Phylogeny for taxonomic relatedness"
author: "Michelle DePrenger-Levin"
date: "2023-01-31"
output: html_document
---
TraitsER

Chamberlain et al 2012 Does phylogeny matter
Albrecht et al 2019 supplemental materials

```{r}
library(brranching)
library(phylocomr)
library(ape)
library(taxize)  
library(phytools)
#V.PhyloMaker2
library("V.PhyloMaker")

library(dplyr)
library(tidyr)

library(ggplot2)
library(ggtree)

library(phylolm)
library(pez)

library(R2jags)
library(runjags)
library(mcmcplots)
library(boot)

load("./Rdatas/plants_data2022-12-10.Rdata")
```

What if i did just group (random effects) of family or genus
```{r}
## Some families have all life histories, a few families fall into one, or primarily into one
table(plants.data@data$Family, plants.data@data$LifeHistory) 
plants.data@data %>%
  group_by(LifeHistory, Family) %>%
  summarise()

table(plants.data@data$Genus, plants.data@data$LifeHistory) 
table(plants.data@data$Family, plants.data@data$LifeHistory) 

data.frame(table(plants.data@data$SpeciesAccepted, plants.data@data$LifeHistory)) %>%
  mutate(Var1 = reorder(Var1, Freq)) %>%
  ggplot(  aes(Var1, Freq, fill = Var2))+
    geom_bar(position = "fill", stat = "identity")

data.frame(table(plants.data@data$Genus, plants.data@data$LifeHistory)) %>%
  mutate(Var1 = reorder(Var1, Freq)) %>%
  ggplot(  aes(Var1, Freq, fill = Var2))+
    geom_bar(position = "fill", stat = "identity")

data.frame(table(plants.data@data$Family, plants.data@data$LifeHistory)) %>%
  mutate(Var1 = reorder(Var1, Freq)) %>%
  ggplot(  aes(Var1, Freq, fill = Var2))+
    geom_bar(position = "fill", stat = "identity")+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  
data.frame(table(plants.data@data$Order, plants.data@data$LifeHistory)) %>%
  mutate(Var1 = reorder(Var1, Freq)) %>%
  ggplot(  aes(Var1, Freq, fill = Var2))+
    geom_bar(position = "fill", stat = "identity")+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

length(unique(plants.data@data$SpeciesAccepted))
length(unique(plants.data@data$Family)) ## 92 not all used in simulations
```


 
```{r}

sp.list <- as.data.frame(plants.data@data[,c("Family","Genus", "SpeciesAccepted")] ) %>%
  mutate(Family = replace(Family, Genus == "Laminaria", "Laminariaceae")) %>% ## large brown alga, oarweed
  mutate(SpeciesAccepted = replace(SpeciesAccepted, SpeciesAccepted == "Silene Ciliata", "Silene ciliata")) %>%
  mutate(Family = replace(Family, Family == "Compositae", "Asteraceae")) %>%
  mutate(Family = replace(Family, Family == "Leguminosae", "Fabaceae")) %>%
  mutate(Family = replace(Family, Family == "Legumiosae", "Fabaceae")) %>% ## misspelling in Kummerowia striata family
  mutate(Family = replace(Family, Family == "Dipsacaceae", "Caprifoliaceae"))%>%
  mutate(Family = replace(Family, Family == "Fzcaceae", "Fucaceae")) %>%
  mutate(Family = replace(Family, Family == "Scropulariaceae", "Scrophulariaceae"))%>%
  group_by(SpeciesAccepted,Genus,Family) %>%
  distinct() %>%
  rename(species = SpeciesAccepted,
            genus = Genus,
            family = Family) 

tree.LH <- phylo.maker(sp.list = sp.list[c("species","genus","family")])
write.tree(tree.LH$scenario.3, paste("./Rdatas/tree_LH",Sys.Date(),".tre",sep=""))
## List to check taxonomy
tree.LH$scenario.3 ## Phylogenetic tree with 314 tips and 302 internal nodes.

tree.LH1 <- phylo.maker(sp.list = sp.list[c("species","genus","family")], scenarios = "S1")
tree.LH1$scenario.1  ## Phylogenetic tree with 314 tips and 301 internal nodes

## Attach LH to each species for node color
## first summarize each species' traits
species.data.sum <- plants.data@data %>%  
  mutate(SpeciesAccepted = replace(SpeciesAccepted, SpeciesAccepted == "Silene Ciliata", "Silene ciliata")) %>%
  mutate(Family = replace(Family, Family == "Compositae", "Asteraceae")) %>%
  mutate(Family = replace(Family, Family == "Leguminosae", "Fabaceae")) %>%
  mutate(Family = replace(Family, Family == "Legumiosae", "Fabaceae")) %>% ## misspelling in Kummerowia striata family
  mutate(Family = replace(Family, Family == "Dipsacaceae", "Caprifoliaceae"))%>%
  mutate(Family = replace(Family, Family == "Fzcaceae", "Fucaceae")) %>%
  mutate(Family = replace(Family, Family == "Scropulariaceae", "Scrophulariaceae"))%>%
  group_by(SpeciesAccepted, Genus, Family) %>%
  summarise(shapeIteromean = mean(shapeItero,na.rm = TRUE),
            adultLongmean = mean(remainingAgeMature,na.rm = TRUE),
            matureAgemean = mean(matureAge,na.rm = TRUE),
            gentimemean = mean(gentime,na.rm = TRUE),
            growthratemean = mean(log(growthrate)))%>%
  mutate(LifeHistory = as.factor(case_when(adultLongmean > 2 & matureAgemean < 3 ~ "FI",
                                 adultLongmean < 3 & matureAgemean < 3 ~ "FS",
                                 adultLongmean > 2 & matureAgemean > 2 ~ "SI",
                                 adultLongmean < 3 & matureAgemean > 2 ~ "SS")),
         Pace = as.factor(case_when(matureAgemean < 3 ~ "fast",
                                    matureAgemean > 2 ~ "slow")))
write.csv(species.data.sum, file = paste("./Rdatas/plants_summarySpeciesList",Sys.Date(),".csv", sep=""))

species.list.join <- tree.LH$species.list %>%
  left_join(species.data.sum, by = c("species" = "SpeciesAccepted")) 
tree.LH$species.list <- species.list.join 
dotTree(tree.LH$scenario.3, x, show.tip.label = TRUE)

tree.LH$species.list[grep("Lactuca",tree.LH$species.list$species),c("LifeHistory","shapeIteromean","matureAgemean","adultLongmean")] # SS
tree.LH$species.list[grep("Pinus",tree.LH$species.list$species),c("species","LifeHistory","shapeIteromean","matureAgemean","adultLongmean")] # Pinus radiata by its matrix claims to be FS, no shapeIero for Pinus maximartinezii, others are slow itero
tree.LH$species.list[grep("Ranunculus",tree.LH$species.list$species),c("species","LifeHistory","shapeIteromean","matureAgemean","adultLongmean")]
levels(factor(tree.LH$species.list$LifeHistory)) ## FI = 1 red, FS = 2 blue, SI = 3 green, SS = 4 orange 

tiporder <- gsub("_"," ",tree.LH$scenario.3$tip.label) ## 314
LHlabel <- unlist(lapply(tiporder, function(x) tree.LH$species.list$LifeHistory[tree.LH$species.list$species == x]))
LHorder <- as.numeric(as.factor(unlist(lapply(tiporder, function(x) tree.LH$species.list$LifeHistory[tree.LH$species.list$species == x]))))


jpeg(paste("./CirclePhylogeny", Sys.Date(), sep = ""), width=200, height=200,units='mm', res=300)
plot(tree.LH$scenario.3, "f", tip.color = 
       c("pink","blue","green","orange")[LHorder], cex = 0.5) 
dev.off()

ggsave(filename = paste("./megaPhylogeny",Sys.Date(),".jpg", sep=""),
  ggplot(tree.LH$scenario.3) +
    geom_tree(layout = "inward_circular")+
    geom_tiplab(size = 2, colour = c("pink","blue","green","orange")[LHorder], hjust = 0.5)+
    theme_void(), 
width=200, height=350,units='mm', dpi=300)

levels(LHlabel)

pairwise.tips.LH <- cophenetic(tree.LH$scenario.3)  ## Cophenetic distance for hierarchical clustering

## <https://cran.r-project.org/web/packages/pez/vignettes/pez-pglmm-overview.pdf> 
nspp <- length(tree.LH$scenario.3$tip.label) ## 314 species
Vphy <- vcv(tree.LH$scenario.3) ## variance-covariance matrix (by default corr = FALSE)
SVphy <- Vphy(det(Vphy)^(1/nspp))## standardize the VCV matrix since more species means larger phylogenetic distances, det is inf

tree.LH$scenario.3$edge
tree.LH$scenario.3$Nnode ## 121 species
plot(tree.LH$scenario.3, edge.width = 2, type = "f", cex.main = .75)
            
```





The picante package
One of the core packages for hypothesis testing in ecophylogenetics is picante (Kembel et al. 2010). picante mainly works with three types of objects: the phylogeny data as a phylo class object; the community presence-absence (binary) or abundance matrix; and a species-trait matrix.

```{r}
## Tree
tree.LH$scenario.3

## community data (or species within a life history?)
comm.LH <- as.matrix(table(species.list.join$LifeHistory, species.list.join$species))
## tips have _ instead of spaces, comm data has spaces, nothing matches
dimnames(comm.LH)[[2]] <- unlist(lapply(dimnames(comm.LH)[[2]], function(x) gsub(" ","_",x)))


## Phylogenetic patters as proxies of life history assignment  
# explain uneven distribution of species throughout life history traits, recently diverged tend to be ecologically similar, similar things to address and have a life history trait  
## Within-assemblage phylogenetic structure   
# Faith’s PD (1992) sums the branch lengths of all co-occurring species in a given site, from the tips to the root of the phylogenetic tree. Higher PD values are given to communities that has more evolutionary divergent taxa and older history, while lower PD values represent assemblages that have taxa with more recent evolutionary history.  


## prune and sort so phylogeny and community or trait data match, exlucdes three algea/lichen
tree.LH.clean <- match.phylo.comm(phy = tree.LH$scenario.3, 
                                  comm = comm.LH)$phy
comm.LH.clean <- match.phylo.comm(phy = tree.LH$scenario.3, 
                                  comm = comm.LH)$comm

par(mar=c(5.1, 4.1, 4.1, 2.1))
## The Phylogenetic sturcutre, PD larger means more evolutionary divergent taxa, older history
##                             PD lower more recent evolutionary history
##                             SR is species richness 
LH.pd <- picante::pd(samp = comm.LH.clean,
                     tree = tree.LH.clean,
                     include.root = FALSE)
cor.test(LH.pd$PD, LH.pd$SR)
plot(LH.pd$PD, 
     LH.pd$SR, 
     xlab = "Phylogenetic Diversity", ylab = "Species Richness in life history trait", 
     pch = 16)

## Randomizations and null models 
# taxa.labels - shuffle taxa across tips
### repeat several times for distribution of null models, calculate standardized effect size of PD for each community using ses.pd()

LH.ses.pd <- ses.pd(samp = comm.LH.clean, 
                        tree = tree.LH.clean, 
                        null.model = "taxa.labels", 
                        runs = 1000)
LH.ses.pd
levels(species.list.join$LifeHistory)
```


# Phylogenetic variance-covariance matrix
## To what degree to closely related taxa fall in the same life history
```{r}
pairwise.tips.LH[1:4,1:4]
cophenDist.LH <- cophenetic.phylo(tree.LH.clean)

# Net relatedness index - mean phylgenetic distances, NRI is obtained by multiplying the standardized effect size mean phylogenetic (pairwise) distances (calculated via a null model of interest) by  −1
# negative phylogenetic overdispersion (within group less closely related than expected)
# positive clustering (within group more closely related than expected from null)

LH.ses.mpd <- ses.mpd(comm.LH.clean, 
                      cophenDist.LH, 
                      null.model = "taxa.labels", 
                      abundance.weighted = FALSE, 
                      runs = 100)

# Calculate NRI  (neg values mean more closely related than null)
LH.NRI <- as.matrix(-1 * ((LH.ses.mpd[,2] - LH.ses.mpd[,3]) /
                            LH.ses.mpd[,4]))
rownames(LH.NRI) <- row.names(LH.ses.mpd)
colnames(LH.NRI) <- "NRI"


```











