TraitsLH

```{r}

rm(list=ls())

library(Rage)
library(popdemo)
library(lefko3)
library(Rcompadre)


library(dplyr)
library(tidyr)
library(purrr)
library(tidyverse)
library(patchwork)

library(popbio)
library(MASS)

```

Borrowed from "Other key life history traits from MPMs" in COMPADRE plant matrix database    
<https://compadre-db.org/Education/article/other-key-life-history-traits-from-mpms>   
```{r}

# Define function
lifeTimeRepEvents <- function(matU, matF, startLife = 1) {
  
  uDim = dim(matU)[1]
  surv = colSums(matU)
  repLifeStages = colSums(matF)
  
  repLifeStages[which(repLifeStages>0)] <- 1
  
  if(missing(matF) | missing(matU)) {stop('matU or matF missing')}
  if(sum(matF, na.rm=T)==0) {stop('matF contains only 0 values')}
  
  # Probability of survival to first reprod event
  Uprime <- matU
  Uprime[, which(repLifeStages==1)] <- 0
  Mprime = matrix(0,2,uDim)
  for (p in 1:uDim[1]) {
    if (repLifeStages[p]==1) {Mprime[2,p] = 1} else {
      Mprime[1,p] = 1-surv[p]
    }
  }
  Bprime = Mprime%*%(ginv(diag(uDim)-Uprime))
  pRep = Bprime[2,startLife]
  
  out = data.frame(pRep = pRep)
  
  # Age at first reproduction (La; Caswell 2001, p 124) AGE OF REPRODUCTIVE MATURITY
  D = diag(c(Bprime[2,]))
  Uprimecond = D%*%Uprime%*%ginv(D)
  expTimeReprod = colSums(ginv(diag(uDim)-Uprimecond))
  La = expTimeReprod[startLife]
  
  out$La = La
  
  # Mean life expectancy conditional on entering the life cycle in the first reproductive stage 
  # longevity after being reproductive given that you made it to be reproductive
  firstRepLifeStage = min(which(repLifeStages==1))
  N = solve(diag(uDim[1])-matU) ## solve finds the inverse (so you can multiply to do division)
  meanRepLifeExpectancy = colSums(N)[firstRepLifeStage]
  
  out$meanRepLifeExpectancy = meanRepLifeExpectancy
  
  # Life expectancy from mean maturity
  remainingMatureLifeExpectancy = colSums(N)[startLife]-La
  
  out$remainingMatureLifeExpectancy = remainingMatureLifeExpectancy
  
  return(out)
}


fi <- "blue"
si <- "brown"
fs <- "chartreuse4"
ss <- "goldenrod3"

```

  
```{r}
## fetch most recent from website, add logical columns to metadata to flag potential problems
compadre <- cdb_fetch("compadre", flag = TRUE)

plants <- compadre %>%
  filter(MatrixSplit == "Divided") %>%
  filter(ProjectionInterval == 1) %>%
  filter(check_irreducible == TRUE) %>%
  subset(MatrixComposite != "Seasonal")

## Plot geographic distribution
ggplot(plants, aes(Lon, Lat))+
  borders(database = "world", fill = "grey80", col = NA) +
  geom_point(alpha = 0.4, color = "blue") +
  scale_x_continuous(breaks = seq(-180, 180, 90), expand = c(0, 0))+
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = "Longitude", y = "Latitude") +
  theme_minimal()


plants.data <- plants %>%
  mutate(Species = SpeciesAccepted) %>%
  mutate(matA = matA(.),
         start_life = mpm_first_active(.)) %>%
  mutate(matU = matU(.),
         start_life = mpm_first_active(.)) %>%  
  ## To get lifeTimeRepEvents to work
  mutate(N_inv = lapply(matU, function(x) diag(dim(x)[1])-x)) %>%  ## solve(diag(uDim[1])-matU) is singular det()==0
  filter(lapply(N_inv, function(x) det(x)) != 0) %>%
  mutate(matF = matF(.), 
         start_life = mpm_first_active(.)) %>%
  mutate(matC = matC(.),
         start_life = mpm_first_active(.)) %>%
  mutate(classes = MatrixClassAuthor(.)) %>%
  mutate(activeclasses = MatrixClassOrganized(.)) %>%
  mutate(matF_all0 = mapply(sum, matF)) %>% # remove matrices where fecundity is all zero
  filter(matF_all0 != 0) %>%
  mutate(seedbank = ifelse(grepl("seedbank|Seedbank|seeds|Seeds",classes), "Yes", "No")) %>%
  mutate(matR = mapply("+", matF, matC)) %>%
  filter(!sapply(matU, function(x) any(colSums(x) > 1))) %>% ## remove any matU where transition over 1
  filter(!is.na(mpm_first_active(.))) %>% # in one case there was an NA
  filter(!sapply(matU, function(x) sum(is.na(matU))>0)) %>% ## need to remove any matUs that are NA
  mutate(gentime = mapply(gen_time, matU, matF)) %>%
  mutate(matureAge = mapply(mature_age, matU, matF, start = start_life)) %>%
  ## instead take the median of checking through start life of each stage and a range of critical survival
  mutate(longevity = mapply(longevity, matU, start = start_life, lx_crit = 0.01)) %>%
  mutate(convage = mapply(qsd_converge, matU, start = mpm_first_active(.))) %>% 
  filter(longevity - convage <= 0) %>%
  mutate(matU = mapply(function(x,y) name_stages(x, names = y, prefix = ""), matU, classes),
         matF = mapply(function(x,y) name_stages(x, names = y, prefix = ""), matF, classes)) %>%
  mutate(netReproRate = mapply(net_repro_rate, matU, matF, method = "start", start = mpm_first_active(.))) %>%
  mutate(exc.class = mapply(grepl, "prop|dorm", activeclasses)) %>%
  mutate(exc.IDs = mapply(which, exc.class)) %>% ## which stages are dormant or propagules
  mutate(lx = mapply(mpm_to_lx, matU, start_life, lx_crit = 0.01)) %>%
  mutate(mx = mapply(mpm_to_mx, matU, matF, start_life, lx_crit = 0.01)) %>%
  mutate(DegreeItero = mapply(entropy_d,lx,mx)) 

## Degree of Iteroparity should be (0,inf) not negative
## shape_rep calculates a shape value of distribution of reproduction over age (-0.5 < S < 0.5)
shapeItero <- c()
for(i in 1:length(plants.data$MatrixID)){
  print(paste(i, plants.data$SpeciesAccepted[i]))
  if(length(plants.data$mx[i][[1]][plants.data$mx[i][[1]]>0])<3){ # must have > 2 nonzero
    shapeItero[i] <- NA
  } else {
  shapeItero[i] <- shape_rep(plants.data$mx[[i]])
}}
plants.data$shapeItero <- shapeItero

## Some matrices are computationally singular and not invertible
for(i in c(1:135,138:903,905:1059,1061:1247,1249,1251:1398,1400:length(plants.data$MatrixID))){
  print(paste(i, plants.data$SpeciesAccepted[i], plants.data$MatrixID[i]))
  lifeTimeRepEvents(plants.data$matU[[i]],plants.data$matF[[i]], plants.data$start_life[[i]])
}

problemMatIDs <- plants.data$MatrixID[c(136:137,904,1060,1248,1250,1399)]

plants.data <- plants.data %>%
  mutate(growthrate = unlist(lapply(matA, popbio::lambda)))

meanAgeMature <- c()
for(i in 1:length(plants.data$MatrixID)){
  print(i)
  if(i %in% c(136:137, 904,1060, 1248,1250, 1399)){ 
    meanAgeMature[i] <- NA
  } else {
    meanAgeMature[i] <- lifeTimeRepEvents(plants.data$matU[[i]],plants.data$matF[[i]], plants.data$start_life[[i]])[2]
  }
}

plants.data$meanAgeMature <- do.call(c, meanAgeMature)

remainingAgeMature <- c()
for(i in 1:length(plants.data$MatrixID)){
  if(i %in% c(136:137, 904,1060, 1248,1250, 1399)){
    remainingAgeMature[i] <- NA
  } else {
    remainingAgeMature[i] <- lifeTimeRepEvents(plants.data$matU[[i]],plants.data$matF[[i]],
                                               plants.data$start_life[[i]])[3]
  }
}
plants.data$remainingAgeMature <- do.call(c, remainingAgeMature)
```


```{r}

plants.data <- plants.data %>%
  mutate(parity = case_when(remainingAgeMature >= 3 ~ "itero",
                            remainingAgeMature < 3 ~ "semel")) %>%
  filter(!is.na(remainingAgeMature)) %>%
  mutate(LifeHistory = case_when(matureAge < 3 & remainingAgeMature < 3 ~ "FS",
                                 matureAge >= 3  & remainingAgeMature < 3 ~ "SS",
                                 matureAge < 3  & remainingAgeMature >= 3 ~ "FI",
                                 matureAge >= 3  & remainingAgeMature >= 3 ~ "SI"))

## Compare to semelparity defined as 1 or fewer years of adult longevity, i.e. age of reproductive maturity followed by death 
plants.dataSI1 <- plants.data %>%
  mutate(parity = case_when(remainingAgeMature >= 3 ~ "itero",
                            remainingAgeMature < 3 ~ "semel")) %>%
  filter(!is.na(remainingAgeMature)) %>%
  mutate(LifeHistory = case_when(matureAge < 3 & remainingAgeMature <= 1 ~ "FS",
                                 matureAge >= 3  & remainingAgeMature <= 1 ~ "SS",
                                 matureAge < 3  & remainingAgeMature > 1 ~ "FI",
                                 matureAge >= 3  & remainingAgeMature > 1 ~ "SI"))

```



### SEEDBANK plants.seedbank
```{r}
plants.seedbank <- plants.data %>%
  filter(sapply(matC, function(x) sum(x)==0)) %>%  ## remove any that call out clonal growth
  filter(seedbank == "Yes" &
           !is.na(LifeHistory)) 
## Fix family errors
plants.data <- plants.data %>%
  mutate(SpeciesAccepted = as.character(SpeciesAccepted)) %>%
  mutate(Family = as.character(Family)) %>%
  mutate(SpeciesAccepted = replace(SpeciesAccepted, SpeciesAccepted == "Silene Ciliata", "Silene ciliata")) %>%
  mutate(Family = replace(Family, Family == "Compositae", "Asteraceae")) %>%
  mutate(Family = replace(Family, Family == "Leguminosae", "Fabaceae")) %>%
  mutate(Family = replace(Family, Family == "Legumiosae", "Fabaceae")) %>% ## misspelling in Kummerowia striata family
  mutate(Family = replace(Family, Family == "Dipsacaceae", "Caprifoliaceae")) %>%
  mutate(Family = replace(Family, Family == "Fzcaceae", "Fucaceae")) %>%
  mutate(Family = replace(Family, Family == "Scropulariaceae", "Scrophulariaceae"))

```

### Collapse each into seed (prop), vegetative (matF == 0), and reproductive (matF > 0)   
       Bromus tectorum with 1: fall seedbank, 5: winter seedbank
       "Solidago altissima" with "old seeds" called an active stage, and then sprout-clones
       "Lupinus lepidus" has reproduction in all the stages, seeds even, might be transposed?

```{r}    

## name stages by if they are prop, vegetative, reproductive and then collapse = list(seed = "prop", veg = "veg", rep = "rep") stages hold names seed, veg, rep

foo <- plants.seedbank %>%
  filter(SpeciesAccepted != "Solidago altissima") %>% ## no seed stage
  mutate(eigenvalues = mapply(function(x) eigen(x)$values, matA)) %>%
  mutate(stages = lapply(activeclasses, function(x) rep("veg", length(x)))) %>%
  mutate(stages = mapply(function(x,y) replace(x, which(colSums(y) > 0), "rep"), stages, matF)) %>%
  mutate(stages = mapply(function(x,y) replace(x, which(y == "prop"), "seed"), stages, activeclasses)) %>%
  mutate(matU = mapply(function(x,y) name_stages(x, names = y, prefix = ""), matU, stages),
         matF = mapply(function(x,y) name_stages(x, names = y, prefix = ""), matF, stages)) %>%
  filter(mapply(function(x) length(which(x == "veg"))>0 , stages)) 

foo2join <- foo@data %>%
  dplyr::select(SpeciesAccepted, Family, Country:Ecoregion, LifeHistory, 
                shapeItero:remainingAgeMature, gentime:longevity)%>%
  distinct() %>%
  arrange(Family, SpeciesAccepted) %>%
  mutate(Analysis = "Seed Stage")

multiMats <- plants.data %>%
  mutate(mat.dims = sapply(matU, function(x) dim(x)[1])) %>%
  group_by(SpeciesAccepted, LifeHistory, mat.dims) %>%
  mutate(SpeciesLH = mapply(function(x,y,z) paste0(x,y,z), SpeciesAccepted,LifeHistory, mat.dims)) %>%
  filter(!is.na(LifeHistory)) %>%
  mutate(NumMats = n(), .groups = "keep") %>%
  filter(NumMats > 2)

multiMats2join <- multiMats@data %>%
  dplyr::select(SpeciesAccepted, Family, Country:Ecoregion, LifeHistory, 
                shapeItero:remainingAgeMature, gentime:longevity)%>%
  distinct() %>%
  arrange(Family, SpeciesAccepted) %>%
  mutate(Analysis = "Environmental stochasticty")

collapsedmats <- list()
for(i in c(1:nrow(foo))){
  print(paste(foo$MatrixID[i],foo$SpeciesAccepted[i],i))
  collapsedmats[[i]] <- mpm_collapse(foo$matU[[i]], foo$matF[[i]], collapse = list(seed = "seed",
                                                                      veg = "veg",
                                                                      rep = "rep"))
  
}

fastitero <- collapsedmats[foo$LifeHistory == "FI"]
fastsemel <- collapsedmats[foo$LifeHistory == "FS"]
slowitero <- collapsedmats[foo$LifeHistory == "SI"]
slowsemel <- collapsedmats[foo$LifeHistory == "SS"]

FImatA <- lapply(fastitero, function(x) x$matA)
FSmatA <- lapply(fastsemel, function(x) x$matA)
SImatA <- lapply(slowitero, function(x) x$matA)
SSmatA <- lapply(slowsemel, function(x) x$matA)

FImA <- data.frame(matrix(do.call(c, FImatA), nrow = length(FImatA), byrow = TRUE))
FSmA <- data.frame(matrix(do.call(c, FSmatA), nrow = length(FSmatA), byrow = TRUE))
SImA <- data.frame(matrix(do.call(c, SImatA), nrow = length(SImatA), byrow = TRUE))
SSmA <- data.frame(matrix(do.call(c, SSmatA), nrow = length(SSmatA), byrow = TRUE))

## Reproduction; birth matrices
FImatF <- lapply(fastitero, function(x) x$matF)
FSmatF <- lapply(fastsemel, function(x) x$matF)
SImatF <- lapply(slowitero, function(x) x$matF)
SSmatF <- lapply(slowsemel, function(x) x$matF)

FImF <- data.frame(matrix(do.call(c, FImatF), nrow = length(FImatF), byrow = TRUE))
FSmF <- data.frame(matrix(do.call(c, FSmatF), nrow = length(FSmatF), byrow = TRUE))
SImF <- data.frame(matrix(do.call(c, SImatF), nrow = length(SImatF), byrow = TRUE))
SSmF <- data.frame(matrix(do.call(c, SSmatF), nrow = length(SSmatF), byrow = TRUE))

FImatU <- lapply(fastitero, function(x) x$matU)
FSmatU <- lapply(fastsemel, function(x) x$matU)
SImatU <- lapply(slowitero, function(x) x$matU)
SSmatU <- lapply(slowsemel, function(x) x$matU)

## All the death matrices (with transitions (0,1]) are estimated by a beta distribution == LHm
FIm <- data.frame(matrix(do.call(c, FImatU), nrow = length(FImatU), byrow = TRUE))
FSm <- data.frame(matrix(do.call(c, FSmatU), nrow = length(FSmatU), byrow = TRUE))
SIm <- data.frame(matrix(do.call(c, SImatU), nrow = length(SImatU), byrow = TRUE))
SSm <- data.frame(matrix(do.call(c, SSmatU), nrow = length(SSmatU), byrow = TRUE))
## stage transitions
trans <- c("s-s","s-v","s-r","v-s","v-v","v-r","r-s","r-v","r-r")
names(FIm) <- trans
names(FSm) <- trans
names(SIm) <- trans
names(SSm) <- trans

names(FImA) <- trans
names(FSmA) <- trans
names(SImA) <- trans
names(SSmA) <- trans
names <- c("Fast iteroparous","Fast semelparous","Slow iteroparous","Slow semelparous")

vitalratecorrelations <- lapply(list(FImA,FSmA,SImA,SSmA), function(LH){
  cortests <- do.call(rbind,apply(combn(1:9, 2), 2, function(x){ 
    ct <- cor.test(LH[,x[1]], LH[,x[2]])
    data.frame(compare = paste(names(LH)[x[1]], names(LH)[x[2]], sep = "_"), 
               cor = ct$estimate, CI95Lower = ct$conf.int[1], CI95Upper = ct$conf.int[2])
                    }))
  cortests
  })

vrc <- lapply(1:4, function(x) data.frame(LifeHistory = names[x], vitalratecorrelations[[x]]))
vrc <- do.call(rbind, vrc)
# vrc <- vrc[complete.cases(vrc),]
corParams <- vrc %>%
  group_by(LifeHistory, compare) %>%
  summarize(corAvg = mean(cor),
         CILAvg = mean(CI95Lower), 
         CIUAvg = mean(CI95Upper), .groups = "keep")

## correlation of survival, growth, retrogression, and reproduction
vitalrates.LH <- plants.seedbank %>%
  # group_by(LifeHistory, SpeciesAccepted, MatrixID) %>%
  mutate(surv = mapply(vr_vec_survival, matU)) %>%
  mutate(grow = mapply(vr_vec_growth, matU, exclude = exc.IDs)) %>%
  mutate(shrink = mapply(vr_vec_shrinkage, matU, exclude = exc.IDs)) %>%
  mutate(stasis = mapply(vr_vec_stasis, matU)) %>%
  mutate(repro = mapply(vr_vec_reproduction, matU, matF))

### Beta distribution on transition rates
LHm <- lapply(1:4, function(i) data.frame(LifeHistory = names[i], list(FIm,FSm,SIm,SSm)[[i]]))
LHm <- do.call(rbind,LHm)

### Poisson distribution on the fecundity
LHmbirth <- lapply(1:4, function(i) data.frame(LifeHistory = names[i], list(FImF,FSmF,SImF,SSmF)[[i]]))
LHmbirth <- do.call(rbind,LHmbirth)
names(LHmbirth) <- c("LifeHistory","ss","sv","sr","vs","vv","vr","rs","rv","rr")
colSums(LHmbirth[,-1])
LHmbirth <- LHmbirth[,c(1,8:9)]

# parameters of beta distribution of seed to seed for fast iteroparous
## should be 16 rows
# LHm[LHm == 0] <- 0.0001
betaParams <- LHm %>%
  dplyr::select(-c(s.r,v.s,r.s)) %>%
  group_by(LifeHistory) %>% ## Reproductive to seed is a poisson and not a beta
  pivot_longer(cols = s.s:r.r, names_to = "transition", values_to = "vitalrate") %>%
  filter(vitalrate != 0) %>% ## beta for [0,1]
  ungroup() %>%
  group_by(LifeHistory, transition) %>%
  summarise(beta1 = EnvStats::ebeta(vitalrate)$parameters[1],
         beta2 = EnvStats::ebeta(vitalrate)$parameters[2], .groups = "keep")  ## throws errors for zero 
         
trans <- unique(betaParams$transition)
p <- seq(0,1,length=100)
betaDist <- betaParams %>%
  group_by(LifeHistory, transition) %>%
  summarise(db = dbeta(p, beta1, beta2), .groups = "keep") 

betaDist <- data.frame(betaDist, p)


vitalrates <- ggplot(betaDist, aes(p, db, color = LifeHistory))+
  geom_line(linewidth = 1)+
  xlim(c(0.00001, 0.99999))+
  facet_wrap(~transition, scales = "free_y",
             labeller = as_labeller(c("s.s" = "Soil Seed Bank",
                                      "s.v" = "Germination",
                                      "v.r" = "Maturation",
                                      "v.v" = "Survival of Vegetative",
                                      "r.r" = "Survival of Reproductive",
                                      "r.v" = "Retrogression")))+
  scale_color_manual("Life history", values = c(fi,fs,si,ss))+
  ylab("PDF")+
  xlab("transition rate")+
  theme_bw()+
  ggtitle("c)")

       
((plants.data %>%
  filter(!is.na(LifeHistory)) %>%
  ggplot( aes(log(matureAge), log(gentime), color = LifeHistory,
              fill = LifeHistory, shape = LifeHistory))+
  geom_point()+
  scale_color_manual("Life history",
                     breaks = c("FI","FS","SI","SS"),
                     labels = c("Fast iteroparous","Fast semelparous",
                                "Slow iteroparous", "Slow Semelparous"),
                     values = c("FI"= fi, "FS"=fs,"SI"=si,"SS"=ss),
                     guide="none")+
  scale_fill_manual("Life history",
                     breaks = c("FI","FS","SI","SS"),
                     labels = c("Fast iteroparous","Fast semelparous",
                                "Slow iteroparous", "Slow Semelparous"),
                     values = c("FI"= fi, "FS"=fs,"SI"=si,"SS"=ss),
                     guide="none")+
  scale_shape_manual("Life history",
                     breaks = c("FI","FS","SI","SS"),
                     labels = c("Fast iteroparous","Fast semelparous",
                                "Slow iteroparous", "Slow Semelparous"),
                     values = c("FI" =4,"FS"=21,"SI"=4,"SS"=21),
                     guide="none")+

  xlab("log(Age of maturity)")+
  ylab("log(Generation time)")+
  theme_bw()+
  geom_abline(slope = 1, intercept = 0)+
    ggtitle("a)"))

+

(  plants.data %>%
  filter(!is.na(LifeHistory)) %>%
  ggplot( aes(LifeHistory, log(growthrate), color = LifeHistory))+
    geom_boxplot()+
    geom_jitter(size = 0.75)+
    # geom_hline(yintercept = 1, col = "grey50", linetype = "dashed")+
    geom_hline(yintercept = 0, col = "grey50", linetype = "dashed")+
    theme_bw()+
  scale_color_manual("Life history",
                     breaks = c("FI","FS","SI","SS"),
                     labels = c("Fast iteroparous","Fast semelparous",
                                "Slow iteroparous", "Slow Semelparous"),
                     values = c("FI"= fi, "FS"=fs,"SI"=si,"SS"=ss))+
    ggtitle("b)")))  + 
 (vitalrates) + plot_layout(widths = c(1.25,1.5,4))


```

