---
title: "Compare asymtotic demo env"
author: "Michelle DePrenger-Levin"
date: "2022-10-25"
output: html_document
---
TraitsER
## Categorical pace and parity
Stage [Seed]  [Seedling]

Compare the impact of founding population size, founding population stage, pace (categorized), and parity (categorized), on extinction risk with asymptotic, demographic, environmental, and demo + envion stochasticity.


```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(patchwork)


load("./Rdatas/plants_data.Rdata")

plants.seedbank <- plants.data %>%
  filter(sapply(matC, function(x) sum(x)==0)) %>%  ## remove any that call out clonal growth
  filter(seedbank == "Yes" &
           !is.na(LifeHistory)) 
```


Asymptotic (control)    
```{r}
Time2Ext_EP <- do.call(rbind,lapply(c(1,10,100,500,1000), function(stpopsz){
  out <- do.call(rbind, lapply(1:10000, function(x){  #length(plants.seedbank$MatrixID) sample with replacement
      ind <- sample(1:length(plants.seedbank$matA),1)
      n <- matrix(rep(0,length(plants.seedbank$activeclasses[[ind]])), ncol = 1, byrow = TRUE) # an empty vector of numbers in stage 
      ## seed start
      ssd <- stable.stage(plants.seedbank$matA[[ind]])
      starthere <- which(ssd == max(ssd["prop" == plants.seedbank$activeclasses[[ind]]])) ## if more than one prop stage, start with the one with the greatest stable stage distribution density
      n[starthere] <- stpopsz

      Time2Ext <- NA
      for(yr in 1:500){
        n_t1 <- floor(plants.seedbank$matA[[ind]] %*% n[,yr])
        if(sum(n_t1)==0| is.na(sum(n_t1))){ 
          Time2Ext <- yr+1
          break
          }
        n <- cbind(n,n_t1)
        }
      seedout <- data.frame(Time2Ext, start = "seed", StPopSz = stpopsz, LH = plants.seedbank$LifeHistory[ind],
                            DegreeItero = plants.seedbank$DegreeItero[ind], AgeMature = plants.seedbank$matureAge[ind],
                            MatrixID = plants.seedbank$MatrixID[ind])
      rm(yr)
      rm(n)
      
      ## Seedling start
      n <- matrix(rep(0,length(plants.seedbank$activeclasses[[ind]])), ncol = 1, byrow = TRUE) # an empty vector of numbers in stage 
      n[plants.seedbank$start_life[[ind]]] <- stpopsz ## the first non-seed stage
      
      Time2Ext <- NA
      for(yr in 1:500){
        n_t1 <- floor(plants.seedbank$matA[[ind]] %*% n[,yr])
        if(sum(n_t1)==0 | is.na(sum(n_t1))){ 
          Time2Ext <- yr+1
          break
          }
        n <- cbind(n,n_t1)
      }  
      seedlingout <- data.frame(Time2Ext, start = "seedling", StPopSz = stpopsz, LH = plants.seedbank$LifeHistory[ind],
                            DegreeItero = plants.seedbank$DegreeItero[ind], AgeMature = plants.seedbank$matureAge[ind],
                            MatrixID = plants.seedbank$MatrixID[ind])
        
      rbind(seedout,seedlingout)
  })) ## end replicates
  out
}))

```

Demographic stochasticity   
```{r}
demostoch <- do.call(rbind,lapply(c(1,10,100,500,1000), function(stpopsz){
  out <- do.call(rbind, lapply(1:10000, function(x){
      ind <- sample(1:length(plants.seedbank$matU),1)
      n <- matrix(rep(0,length(plants.seedbank$activeclasses[[ind]])), nrow = 1, byrow = TRUE)
      
      #Seed start
        ssd <- stable.stage(plants.seedbank$matA[[ind]])
        ## if more than one prop stage, start with the one with the greatest stable stage distribution density
        starthere <- which(ssd == max(ssd["prop" == plants.seedbank$activeclasses[[ind]]])) 
        n[starthere] <- stpopsz
      
        colnames(n) <- colnames(plants.seedbank$matU[[ind]])
        
        Time2Ext <- NA
      # Each row is a year for number in each stage
      for(yr in 1:100){
        grow_surv <- colSums(apply(plants.seedbank$matU[[ind]],1,function(m_row) {
          sapply(1:length(plants.seedbank$activeclasses[[ind]]), function(x){
            rbinom(1, size = n[yr,x], prob = m_row[x])
            })
          }))
        fert <- colSums(apply(plants.seedbank$matF[[ind]], 1, function(m_row) {
          sapply(1:length(plants.seedbank$activeclasses[[ind]]), function(x){
            rpois(1, m_row[x]*n[yr,x])
          })
        }))
        if(sum(grow_surv+fert) == 0 | is.na(sum(grow_surv+fert))){ 
          Time2Ext <- yr
          break
          }
        n <- rbind(n,grow_surv + fert)
      }
      seedout <- data.frame(Time2Ext, start = "seed", StPopSz = stpopsz, LH = plants.seedbank$LifeHistory[ind],
                      DegreeItero = plants.seedbank$DegreeItero[ind], AgeMature = plants.seedbank$matureAge[ind],
                      MatrixID = plants.seedbank$MatrixID[ind])
      
      
      ## Seedling start
        rm(n)
        n <- matrix(rep(0,length(plants.seedbank$activeclasses[[ind]])), nrow = 1, byrow = TRUE)
        n[plants.seedbank$start_life[[ind]]] <- stpopsz ## the first non-seed stage
        Time2Ext <- NA
        
        for(yr in 1:100){
        grow_surv <- colSums(apply(plants.seedbank$matU[[ind]],1,function(m_row) {
          sapply(1:length(plants.seedbank$activeclasses[[ind]]), function(x){
            rbinom(1, size = n[yr,x], prob = m_row[x])
            })
          }))
        fert <- colSums(apply(plants.seedbank$matF[[ind]], 1, function(m_row) {
          sapply(1:length(plants.seedbank$activeclasses[[ind]]), function(x){
            rpois(1, m_row[x]*n[yr,x])
          })
        }))
        if(sum(grow_surv+fert | is.na(sum(grow_surv+fert))) == 0){ 
          Time2Ext <- yr
          break
          }
        n <- rbind(n,grow_surv + fert)
      }
      seedlingout <- data.frame(Time2Ext, start = "seedling", StPopSz = stpopsz, LH = plants.seedbank$LifeHistory[ind],
                      DegreeItero = plants.seedbank$DegreeItero[ind], AgeMature = plants.seedbank$matureAge[ind],
                      MatrixID = plants.seedbank$MatrixID[ind])
       rbind(seedout,seedlingout)
       }))
  out
}))
```


Environmental stochasticity   
```{r}

## only species represented by 3 or more matrices; 20 species
multiMats <- plants.data %>%
  mutate(mat.dims = sapply(matU, function(x) dim(x)[1])) %>%
  group_by(SpeciesAccepted, LifeHistory, mat.dims) %>%
  mutate(SpeciesLH = mapply(function(x,y,z) paste0(x,y,z), SpeciesAccepted,LifeHistory, mat.dims)) %>%
  filter(!is.na(LifeHistory)) %>%
  mutate(NumMats = n(), .groups = "keep") %>%
  filter(NumMats > 2)

multiMats@data %>%
  dplyr::select(SpeciesAccepted, Family, Country:Ecoregion)%>%
  distinct() %>%
  arrange(Family, SpeciesAccepted)

data.frame(multiMats@data[!duplicated("SpeciesAccepted"),c("SpeciesAccepted", "Family","Country","Ecoregion","CensusType")])

table(multiMats$LifeHistory)

environ_portfolio <- do.call(rbind,lapply(c(1,10,100,500,1000), function(stpopsz){
  out <- do.call(rbind, lapply(1:10000, function(x){
    sp <- sample(multiMats$SpeciesLH,1)
    ## A vector the length of stages
    n <- matrix(rep(0,length(multiMats$activeclasses[multiMats$SpeciesLH == sp][[1]]), ncol= 1, byrow = TRUE))
    
    ## seed start
    ssd <- stable.stage(mean(multiMats$matA[multiMats$SpeciesLH == sp]))
    starthere <- which(ssd == max(ssd[grepl("prop", multiMats$activeclasses[multiMats$SpeciesLH == sp][[1]])]))
    
    n[starthere] <- stpopsz
    
    ### Build multistate model Caswell 2001 pg 62:67 with no 'dispersal' among environments
    ## block diagonal
    # multistatemodel <- Matrix::bdiag(multiMats$matA[multiMats$SpeciesAccepted == sp])
    
    Time2Ext <- NA
    for(yr in 1:100){ ## cycle through each year
      matA <- sample(multiMats$matA[multiMats$SpeciesLH == sp],1)[[1]]
      n_t1 <- floor(matA %*% n[,yr])
      n <- cbind(n,n_t1)
      if(sum(n_t1) == 0){
        Time2Ext <- yr
        break
      }
    }
    seedout <- data.frame(Time2Ext, start = "seed", StPopSz = stpopsz,
                          LH = multiMats$LifeHistory[multiMats$SpeciesLH == sp][1],
                          DegreeItero = mean(multiMats$DegreeItero[multiMats$SpeciesLH == sp]), 
                          ShapeItero = mean(multiMats$shapeItero[multiMats$SpeciesLH == sp]),
                          AgeMature = mean(multiMats$matureAge[multiMats$SpeciesLH == sp]),
                          MatrixID = paste(multiMats$MatrixID[multiMats$SpeciesLH == sp],
                                           collapse = "_"))
    rm(yr)
    rm(n)
    
    ## Seedling start
    n <- matrix(rep(0,length(multiMats$activeclasses[multiMats$SpeciesLH == sp][[1]]), ncol= 1, byrow = TRUE))

    n[multiMats$start_life[multiMats$SpeciesLH == sp][1]] <- stpopsz
    
    Time2Ext <- NA
    for(yr in 1:500){ ## cycle through each year
      matA <- sample(multiMats$matA[multiMats$SpeciesLH == sp],1)[[1]]
      n_t1 <- floor(matA %*% n[,yr])
      n <- cbind(n,n_t1)
      if(sum(n_t1) == 0){
        Time2Ext <- yr
        break
      }
    }
    seedlingout <- data.frame(Time2Ext, start = "seedling", StPopSz = stpopsz,
                          LH = multiMats$LifeHistory[multiMats$SpeciesLH == sp][1],
                          DegreeItero = mean(multiMats$DegreeItero[multiMats$SpeciesLH == sp]), 
                          ShapeItero = mean(multiMats$shapeItero[multiMats$SpeciesLH == sp]),
                          AgeMature = mean(multiMats$matureAge[multiMats$SpeciesLH == sp]),
                          MatrixID = paste(multiMats$MatrixID[multiMats$SpeciesLH == sp],
                                           collapse = "_"))
    
    rbind(seedout, seedlingout)
  }))
  out
}))

```



Demographic and environmental stochasticity    
```{r}
## only species represented by 3 or more matrices; 20 species
multiMats <- plants.seedbank %>%
  mutate(mat.dims = sapply(matU, function(x) dim(x)[1])) %>%
  group_by(SpeciesAccepted, LifeHistory, mat.dims) %>%
  mutate(SpeciesLH = mapply(function(x,y,z) paste0(x,y,z), SpeciesAccepted,LifeHistory, mat.dims)) %>%
  filter(!is.na(LifeHistory)) %>%
  mutate(NumMats = n(), .groups = "keep") %>%
  filter(NumMats > 2)

table(multiMats$LifeHistory)

environ_portfolio_demo <- do.call(rbind,lapply(c(1,10,100,500,1000), function(stpopsz){
  out <- do.call(rbind, lapply(1:10000, function(x){
    sp <- sample(multiMats$SpeciesLH,1)
    n <- matrix(rep(0,multiMats$mat.dims[multiMats$SpeciesLH == sp][1], ncol= 1, byrow = TRUE))
    
    ## seed start
    ssd <- stable.stage(mean(multiMats$matA[multiMats$SpeciesLH == sp]))
    starthere <- which(ssd == max(ssd[grepl("prop", multiMats$activeclasses[multiMats$SpeciesLH == sp][[1]])]))
    n[starthere] <- stpopsz
    
    ### Build multistate model Caswell 2001 pg 62:67 with no 'dispersal' among environments
      ## block diagonal
     # deathBlock <- Matrix::bdiag(multiMats$matU[multiMats$SpeciesAccepted == sp & multiMats$LifeHistory == lh])
     # birthBlock <- Matrix::bdiag(multiMats$matF[multiMats$SpeciesAccepted == sp & multiMats$LifeHistory == lh])
      
      Time2Ext <- NA
      for(yr in 1:100){ ## cycle through each year
        print(paste(yr, sp))
        ## Sum rows for how many individuals survived (status or grew into that stage)
        ## by row 
        ## by col enter the number of individuals that would be in that cell and prob of that transition
        ## n is a column per year
        ind <- sample(1:length(multiMats$matU[multiMats$SpeciesLH == sp]),1)
        matU <- multiMats$matU[multiMats$SpeciesLH == sp][[ind]]
        matF <- multiMats$matF[multiMats$SpeciesLH == sp][[ind]]
      if(nrow(n)>1){
        grow_surv <- colSums(apply(matU,1,function(m_row) {
            sapply(1:multiMats$mat.dims[multiMats$SpeciesLH == sp][[1]], function(x){
              rbinom(1, size = n[x,yr], prob = m_row[x])
              })
            }))
          fert <- colSums(apply(matF, 1, function(m_row) {
            sapply(1:multiMats$mat.dims[multiMats$SpeciesLH == sp][[1]], function(x){
              min(rpois(1, m_row[x]*n[x,yr]),1e9) ## NAs if value too large
            })
          }))
          } else {
            grow_surv <- rbinom(1, size = n[,yr], prob = matU)
            fert <- rpois(1, matF*n[,yr])
          }
        
          if(sum(grow_surv+fert) == 0){  
            Time2Ext <- yr
            break
            }
          n <- cbind(n,grow_surv + fert)
      }
      
      seedout <- data.frame(Time2Ext, start = "seed", StPopSz = stpopsz, 
                            LH = multiMats$LifeHistory[multiMats$SpeciesLH == sp][1],
                            DegreeItero = mean(multiMats$DegreeItero[multiMats$SpeciesLH == sp]), 
                            ShapeItero = mean(multiMats$shapeItero[multiMats$SpeciesLH == sp]),
                            AgeMature = mean(multiMats$matureAge[multiMats$SpeciesLH == sp]),
                            MatrixID = paste(multiMats$MatrixID[multiMats$SpeciesLH == sp],
                                             collapse = "_"))
      rm(yr)
      rm(n)
      
      ## Seedling start
      n <- matrix(rep(0,multiMats$mat.dims[multiMats$SpeciesLH == sp][1], ncol= 1, byrow = TRUE))
    
      n[multiMats$start_life[multiMats$SpeciesLH == sp][[1]]] <- stpopsz
      
      Time2Ext <- NA
      for(yr in 1:100){
        ind <- sample(1:length(multiMats$matU[multiMats$SpeciesLH == sp]),1)
        matU <- multiMats$matU[multiMats$SpeciesLH == sp][[ind]]
        matF <- multiMats$matF[multiMats$SpeciesLH == sp][[ind]]
      if(nrow(n)>1){
        grow_surv <- colSums(apply(matU,1,function(m_row) {
            sapply(1:multiMats$mat.dims[multiMats$SpeciesLH == sp][[1]], function(x){
              rbinom(1, size = n[x,yr], prob = m_row[x])
              })
            }))
        fert <- colSums(apply(matF, 1, function(m_row) {
          sapply(1:multiMats$mat.dims[multiMats$SpeciesLH == sp][[1]], function(x){
            min(rpois(1, m_row[x]*n[x,yr]),1e9) ## NAs if value too large
            })
          }))
          } else {
            grow_surv <- rbinom(1, size = n[,yr], prob = matU)
            fert <- rpois(1, matF*n[,yr])
          }
        if(sum(grow_surv+fert) == 0){ 
          Time2Ext <- yr
          break
          }
        n <- cbind(n,grow_surv + fert)
        }
      
      seedlingout <- data.frame(Time2Ext, start = "seedling", StPopSz = stpopsz, 
                                LH = multiMats$LifeHistory[multiMats$SpeciesLH == sp][1],
                                DegreeItero = mean(multiMats$DegreeItero[multiMats$SpeciesLH == sp]), 
                                ShapeItero = mean(multiMats$shapeItero[multiMats$SpeciesLH == sp]), 
                                AgeMature = mean(multiMats$matureAge[multiMats$SpeciesLH == sp]),
                                MatrixID = paste(multiMats$MatrixID[multiMats$SpeciesLH == sp],
                                                 collapse = "_"))
      
      rbind(seedout, seedlingout)
    })) # end out
    out
    }))



```

